---
title: "Okanagan Civic Pulse"
---

```{r setup}
#| include: false
# Rendered from dashboard/, so repo paths are relative to the parent.
for (f in c("theme_okcp.R", "election.R", "taxonomy.R", "findings.R"))
  source(file.path("..", "R", f))
suppressMessages({library(DBI); library(dplyr); library(ggplot2); library(scales)})

# Read the DE-TEXTED serve copy, never the live ingest DB: DuckDB is
# single-writer, so a render during a crawl would fail on a lock, and the serve
# copy additionally contains no comment text and no handles.
serve <- file.path("..", "db", "serve.duckdb")
HAVE_DATA <- file.exists(serve)
if (HAVE_DATA) {
  con <- dbConnect(duckdb::duckdb(), serve, read_only = TRUE)
  issue_daily <- dbGetQuery(con, "SELECT * FROM issue_daily")
  post_issues <- dbGetQuery(con, "SELECT * FROM post_issues")
  posts_meta  <- dbGetQuery(con, "SELECT * FROM posts_meta")
  edges       <- dbGetQuery(con, "SELECT * FROM edges")
  nodes       <- dbGetQuery(con, "SELECT * FROM nodes")
  scrape_log  <- dbGetQuery(con, "SELECT * FROM scrape_log ORDER BY run_at DESC")
  sentiment   <- tryCatch(dbGetQuery(con, "SELECT * FROM sentiment"), error = function(e) data.frame())
  src_daily   <- tryCatch(dbGetQuery(con, "SELECT * FROM source_daily"), error = function(e) data.frame())
  sources_tbl <- tryCatch(dbGetQuery(con, "SELECT * FROM sources"), error = function(e) data.frame())
  # forums carries `region`, which generate_findings() needs to separate the
  # Okanagan forums from the out-of-region ones. Without it the scope-split
  # headline silently pools all nine and reports a collection decision as a
  # finding.
  forums      <- tryCatch(dbGetQuery(con, "SELECT * FROM forums"), error = function(e) data.frame())
  ergm_fit    <- tryCatch(dbGetQuery(con, "SELECT * FROM ergm_fit"), error = function(e) data.frame())
  net_tl      <- tryCatch(dbGetQuery(con, "SELECT * FROM net_timeline"), error = function(e) data.frame())
  turnover    <- tryCatch(dbGetQuery(con, "SELECT * FROM actor_turnover"), error = function(e) data.frame())
  dbDisconnect(con, shutdown = TRUE)
} else {
  issue_daily <- post_issues <- posts_meta <- edges <- nodes <- scrape_log <-
    sentiment <- src_daily <- sources_tbl <- ergm_fit <- net_tl <- turnover <-
    forums <- data.frame()
}

LOCAL_SCOPES <- c("local", "ballot", "shared")

issue_label <- function(code) {
  ifelse(code %in% ISSUES$code, ISSUES$label[match(code, ISSUES$code)], code)
}
n_or <- function(x, alt = 0) if (length(x) && !is.na(x)) x else alt
fresh_days <- if (nrow(posts_meta))
  as.integer(Sys.Date() - as.Date(max(posts_meta$posted_at, na.rm = TRUE))) else NA_integer_
```

# Overview

## Row {height="22%"}

```{r}
#| content: valuebox
#| title: "Days to voting day"
list(icon = "calendar-event", color = "primary", value = days_to_vote())
```

```{r}
#| content: valuebox
#| title: "Election phase"
list(icon = "flag", color = "light", value = election_phase())
```

```{r}
#| content: valuebox
#| title: "Comments analysed"
list(icon = "chat-left-text", color = "light",
     value = format(nrow(posts_meta), big.mark = ","))
```

```{r}
#| content: valuebox
#| title: "Locally actionable"
local_share <- if (nrow(post_issues)) {
  pl <- post_issues |> group_by(post_id) |>
    summarise(loc = any(scope %in% c("local","ballot","shared")), .groups = "drop")
  sprintf("%.0f%%", 100 * mean(pl$loc))
} else "—"
list(icon = "geo-alt", color = "primary", value = local_share)
```

## Row {height="78%"}

### Column

```{r}
#| title: "Where the discourse actually lands"
if (!nrow(post_issues)) {
  cat("No model-coded comments yet. Run the classification pass.")
} else {
  d <- post_issues |> filter(issue_code != "none") |>
    count(scope, name = "n") |>
    mutate(scope = factor(scope, levels = names(SCOPE_COLOURS)))
  ggplot(d, aes(reorder(scope, n), n, fill = scope)) +
    geom_col(width = 0.7) +
    geom_text(aes(label = comma(n)), hjust = -0.15, size = pt(12), colour = BRAND$ink) +
    scale_fill_manual(values = SCOPE_COLOURS, guide = "none") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.18)), labels = comma) +
    coord_flip() +
    labs(x = NULL, y = "Issue mentions",
         title = "Scope of governance issues raised",
         subtitle = "local / ballot / shared are actionable by bodies elected 2026-10-17",
         caption = cap("Aggregate counts from model-coded comments. Keyword-sieve rows excluded.")) +
    theme_okcp()
}
```

### Column

```{r}
#| title: "Top issues"
if (!nrow(post_issues)) {
  cat("No model-coded comments yet.")
} else {
  d <- post_issues |> filter(issue_code != "none") |>
    count(issue_code, scope, name = "n") |>
    slice_max(n, n = 12) |>
    mutate(label = issue_label(issue_code))
  ggplot(d, aes(reorder(label, n), n, fill = scope)) +
    geom_col(width = 0.72) +
    scale_fill_manual(values = SCOPE_COLOURS, name = "Scope") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.1)), labels = comma) +
    coord_flip() +
    labs(x = NULL, y = "Mentions", title = "Most-raised issues") +
    theme_okcp() + theme(legend.position = "bottom")
}
```

# Timeline

## Row {height="60%"}

```{r}
#| title: "Issue salience against the election calendar"
if (!nrow(issue_daily)) {
  cat("No daily rollup yet — run rebuild_issue_daily().")
} else {
  bands <- election_bands()
  top <- issue_daily |> filter(scope %in% c("local","ballot","shared")) |>
    count(issue_code, wt = n_posts, name = "tot") |> slice_max(tot, n = 6) |> pull(issue_code)
  d <- issue_daily |> filter(issue_code %in% top) |>
    group_by(day, issue_code) |>
    summarise(n = sum(n_posts), .groups = "drop") |>
    mutate(day = as.Date(day), label = issue_label(issue_code))

  # Bands are drawn first so the series sit on top of them, and the x range is
  # forced out to voting day so the run-up is visible even before it fills in.
  xlim <- range(c(d$day, ELECTION$pre_campaign_from, ELECTION$voting_day))
  ggplot() +
    geom_rect(data = bands,
              aes(xmin = from, xmax = to, ymin = -Inf, ymax = Inf, fill = phase),
              alpha = 0.14) +
    scale_fill_manual(values = c("Pre-campaign" = BRAND$accent,
                                 "Nominations"  = BRAND$muted,
                                 "Campaign"     = BRAND$warm), name = "Election phase") +
    geom_vline(xintercept = ELECTION$voting_day, linetype = "22", colour = BRAND$warm) +
    annotate("text", x = ELECTION$voting_day, y = Inf, label = "  Voting day",
             hjust = 0, vjust = 1.6, size = pt(12), colour = BRAND$warm) +
    geom_line(data = d, aes(day, n, colour = label), linewidth = 0.8) +
    scale_x_date(limits = xlim, date_labels = "%b %d") +
    scale_colour_brewer(palette = "Dark2", name = NULL) +
    labs(x = NULL, y = "Comments per day",
         title = "Locally actionable issues over time",
         subtitle = sprintf("Voting day 2026-10-17 · currently %s", election_phase()),
         caption = cap("Bands: Elections BC pre-campaign (Jul 20 - Sep 18), nominations (Sep 1-11), campaign (Sep 19 - Oct 17).")) +
    theme_okcp() + theme(legend.position = "bottom")
}
```

## Row {height="40%"}

```{r}
#| title: "Local vs out-of-scope share, weekly"
if (!nrow(issue_daily)) {
  cat("No daily rollup yet.")
} else {
  d <- issue_daily |>
    mutate(week = as.Date(cut(as.Date(day), "week")),
           band = ifelse(scope %in% c("local","ballot","shared"),
                         "Locally actionable", "Provincial / federal")) |>
    group_by(week, band) |> summarise(n = sum(n_posts), .groups = "drop")
  ggplot(d, aes(week, n, fill = band)) +
    geom_col(position = "fill", width = 6) +
    scale_fill_manual(values = c("Locally actionable" = BRAND$accent,
                                 "Provincial / federal" = BRAND$warm), name = NULL) +
    scale_y_continuous(labels = percent, expand = c(0, 0)) +
    scale_x_date(date_labels = "%b %d") +
    labs(x = NULL, y = "Share of issue mentions",
         title = "How much of the anger is a council's to answer?") +
    theme_okcp() + theme(legend.position = "bottom")
}
```

# Issues

```{r}
#| title: "Issue register"
if (!nrow(post_issues)) {
  cat("No model-coded comments yet.")
} else {
  tab <- post_issues |> filter(issue_code != "none") |>
    group_by(issue_code, scope) |>
    summarise(mentions = n(), posts = n_distinct(post_id),
              mean_salience = round(mean(salience, na.rm = TRUE), 2),
              mean_confidence = round(mean(confidence, na.rm = TRUE), 2),
              .groups = "drop") |>
    mutate(Issue = issue_label(issue_code)) |>
    arrange(desc(mentions)) |>
    select(Issue, Scope = scope, Mentions = mentions, Posts = posts,
           `Mean salience` = mean_salience, `Mean confidence` = mean_confidence)
  knitr::kable(tab)
}
```

# Jurisdictions

```{r}
#| title: "Locally actionable mentions by community"
if (!nrow(post_issues)) {
  cat("No model-coded comments yet.")
} else {
  d <- post_issues |>
    filter(!is.na(jurisdiction), jurisdiction != "unspecified",
           scope %in% c("local","ballot","shared")) |>
    count(jurisdiction, name = "n") |> slice_max(n, n = 15) |>
    mutate(label = ifelse(jurisdiction %in% JURISDICTIONS$code,
                          JURISDICTIONS$label[match(jurisdiction, JURISDICTIONS$code)],
                          jurisdiction))
  if (!nrow(d)) cat("No jurisdiction-tagged local issues yet.") else
  ggplot(d, aes(reorder(label, n), n)) +
    geom_col(fill = BRAND$accent, width = 0.7) +
    geom_text(aes(label = comma(n)), hjust = -0.2, size = pt(12), colour = BRAND$ink) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.18)), labels = comma) +
    coord_flip() +
    labs(x = NULL, y = "Locally actionable mentions",
         title = "Which communities the actionable discourse is about") +
    theme_okcp()
}
```

# Methods

## Row

::: {.card title="What this measures"}

Comments on Castanet's public phpBB forums are coded against a codebook of BC
local-government competencies drawn from the Community Charter and Local
Government Act. The central judgement is **scope**: whether an issue is one a
municipal council or regional board can act on, or a provincial/federal matter
being directed at city hall.

`ballot` is deliberately separate from `local` — school trustees and regional
district electoral area directors are not municipal, but their seats are on the
same 2026-10-17 ballot.

:::

::: {.card title="How comments are selected and coded"}

A keyword sieve decides which comments are worth a model call. **The sieve is a
recall filter, not a finding** — every figure on this dashboard excludes sieve
rows and uses model codes only.

The gate requires at least one locally-actionable candidate code. Comments
whose only matches are federal or provincial (foreign affairs, party politics,
courts, provincial highways) are not sent to the model, which cut eligible
comments by 28% on the pilot corpus.

Bulk coding runs on a small model; a stratified sample is re-coded by a larger
one to **measure** agreement. The audit sample deliberately includes comments
the gate *skipped*, so the false-negative rate is estimated rather than assumed.

:::

::: {.card title="Provenance and limits"}

```{r}
if (nrow(scrape_log)) {
  knitr::kable(head(scrape_log[, c("run_at","mode","threads_seen","posts_new","errors")], 5))
} else cat("No crawl log available.\n")
```

- **Post-moderation.** The captured corpus is what survived moderation; removed
  comments and closed comment sections are invisible here.
- **Not a survey.** Forum commenters are not a representative sample of
  residents. Volume is a measure of discourse, not of public opinion.
- **Concentration matters.** Volume is reported next to a Herfindahl index on
  posts-per-actor, because a handful of prolific posters can manufacture the
  appearance of a groundswell.
- **No comment text or handles** appear here or in the underlying file this
  page reads.

:::

```{r}
#| output: asis
cat(sprintf("<div class='provenance'>Corpus freshness: most recent comment %s.<br>%s &middot; %s</div>",
            if (is.na(fresh_days)) "unknown" else
              sprintf("%d day%s old", fresh_days, ifelse(fresh_days == 1, "", "s")),
            BRAND$org, BRAND$product))
```

# Community pulse

## Row {height="55%"}

```{r}
#| title: "Top 5 issues — everything people are talking about"
# Deliberately UNFILTERED by scope. The rest of this dashboard is about what a
# council can act on; this panel answers the different and equally useful
# question of what the community is actually preoccupied with, including the
# things city hall has no lever for.
if (!nrow(post_issues)) {
  cat("No model-coded comments yet.")
} else {
  d <- post_issues |> filter(issue_code != "none") |>
    count(issue_code, scope, name = "n") |>
    slice_max(n, n = 5) |>
    mutate(label = issue_label(issue_code),
           band = ifelse(scope %in% LOCAL_SCOPES, "Council can act", "Not council's to fix"))
  ggplot(d, aes(reorder(label, n), n, fill = band)) +
    geom_col(width = 0.68) +
    geom_text(aes(label = comma(n)), hjust = -0.15, size = pt(12), colour = BRAND$ink) +
    scale_fill_manual(values = c("Council can act" = BRAND$accent,
                                 "Not council's to fix" = BRAND$warm), name = NULL) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.2)), labels = comma) +
    coord_flip() +
    labs(x = NULL, y = "Mentions",
         title = "What the community is talking about",
         subtitle = "All issues, regardless of which government owns them",
         caption = cap("Colour marks whether a body elected on 2026-10-17 has authority over the issue.")) +
    theme_okcp() + theme(legend.position = "bottom")
}
```

## Row {height="45%"}

```{r}
#| title: "Pulse summary"
if (!nrow(post_issues)) {
  cat("No model-coded comments yet.")
} else {
  top5 <- post_issues |> filter(issue_code != "none") |>
    count(issue_code, scope, name = "mentions") |> slice_max(mentions, n = 5)
  posts_n <- n_distinct(post_issues$post_id)
  tab <- top5 |>
    mutate(Issue = issue_label(issue_code),
           Owner = ifelse(scope %in% LOCAL_SCOPES,
                          "Elected 2026-10-17", "Provincial / federal"),
           Share = percent(mentions / sum(top5$mentions), accuracy = 1)) |>
    select(Issue, Scope = scope, Owner, Mentions = mentions, `Share of top 5` = Share)
  knitr::kable(tab)
}
```

# Local government filter

```{r}
#| title: "Issues a council, regional board or school board can act on"
# The filter itself: a client-side searchable table. Static HTML, no server —
# DataTables does the filtering in the browser, so this works on free CDN
# hosting with no backend at all.
if (!nrow(post_issues)) {
  cat("No model-coded comments yet.")
} else {
  d <- post_issues |>
    filter(issue_code != "none", scope %in% LOCAL_SCOPES) |>
    group_by(issue_code, scope, jurisdiction) |>
    summarise(mentions = n(), posts = n_distinct(post_id),
              salience = round(mean(salience, na.rm = TRUE), 2),
              confidence = round(mean(confidence, na.rm = TRUE), 2),
              opposed = percent(mean(stance == "oppose", na.rm = TRUE), accuracy = 1),
              .groups = "drop") |>
    mutate(Issue = issue_label(issue_code),
           Jurisdiction = ifelse(is.na(jurisdiction) | jurisdiction == "unspecified",
                                 "(unspecified)", jurisdiction)) |>
    arrange(desc(mentions)) |>
    select(Issue, Scope = scope, Jurisdiction, Mentions = mentions, Posts = posts,
           Salience = salience, Confidence = confidence, Opposed = opposed)
  DT::datatable(d, rownames = FALSE, filter = "top",
                extensions = "Buttons",
                options = list(pageLength = 15, dom = "Bfrtip",
                               buttons = list("copy", "csv"),
                               order = list(list(3, "desc"))))
}
```

::: {.provenance}
`Scope` distinguishes **local** (direct municipal authority), **ballot** (school
trustees and regional district electoral area directors — not municipal, but on
the same ballot), and **shared** (a real local role, province holds the main
levers). Filter boxes above each column are live; the CSV button exports what
you have filtered to.
:::

# Emotion

## Row {height="55%"}

```{r}
#| title: "Emotional register by issue (per 100 words)"
# NRC returns RAW COUNTS which scale with comment length. On the pilot corpus,
# "oppose" comments looked angrier than neutral ones purely because they are
# 63% longer; normalising per 100 words reversed the comparison. Everything
# here is normalised.
if (!nrow(sentiment) || !nrow(post_issues)) {
  cat("No sentiment scores yet — run score_sentiment().")
} else {
  top <- post_issues |> filter(issue_code != "none") |>
    count(issue_code, name = "n") |> slice_max(n, n = 8) |> pull(issue_code)
  d <- post_issues |> filter(issue_code %in% top) |>
    inner_join(sentiment, by = "post_id") |>
    mutate(across(c(anger, fear, trust, sadness),
                  ~ 100 * .x / pmax(n_words, 1))) |>
    group_by(issue_code, scope) |>
    summarise(across(c(anger, fear, trust, sadness), ~ mean(.x, na.rm = TRUE)),
              n = n(), .groups = "drop") |>
    tidyr::pivot_longer(c(anger, fear, trust, sadness),
                        names_to = "emotion", values_to = "rate") |>
    mutate(label = issue_label(issue_code),
           emotion = factor(emotion, c("anger","fear","sadness","trust")))
  ggplot(d, aes(reorder(label, rate), rate, fill = emotion)) +
    geom_col(position = "dodge", width = 0.75) +
    scale_fill_manual(values = c(anger = "#C05621", fear = "#B08900",
                                 sadness = "#6B7A82", trust = "#1F7A8C"), name = NULL) +
    coord_flip() +
    labs(x = NULL, y = "NRC emotion words per 100 words",
         title = "Which issues carry which emotions",
         caption = cap("Length-normalised. Raw NRC counts scale with comment length and are misleading.")) +
    theme_okcp() + theme(legend.position = "bottom")
}
```

## Row {height="45%"}

```{r}
#| title: "The signal: emotion by who owns the issue"
if (!nrow(sentiment) || !nrow(post_issues)) {
  cat("No sentiment scores yet.")
} else {
  d <- post_issues |> filter(issue_code != "none") |>
    inner_join(sentiment, by = "post_id") |>
    mutate(band = ifelse(scope %in% LOCAL_SCOPES,
                         "Council can act", "Not council's to fix"),
           across(c(anger, fear, trust), ~ 100 * .x / pmax(n_words, 1))) |>
    group_by(band) |>
    summarise(across(c(anger, fear, trust), ~ mean(.x, na.rm = TRUE)),
              words = mean(n_words, na.rm = TRUE), n = n(), .groups = "drop") |>
    tidyr::pivot_longer(c(anger, fear, trust), names_to = "emotion", values_to = "rate")
  ggplot(d, aes(emotion, rate, fill = band)) +
    geom_col(position = "dodge", width = 0.65) +
    geom_text(aes(label = sprintf("%.2f", rate)),
              position = position_dodge(width = 0.65), vjust = -0.4,
              size = pt(12), colour = BRAND$ink) +
    scale_fill_manual(values = c("Council can act" = BRAND$accent,
                                 "Not council's to fix" = BRAND$warm), name = NULL) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
    labs(x = NULL, y = "Words per 100",
         title = "The angriest discourse is about what a council cannot fix",
         caption = cap("Length-normalised, so this is not an artifact of comment length.")) +
    theme_okcp() + theme(legend.position = "bottom")
}
```

# Sources

## Row {height="50%"}

```{r}
#| title: "Is each source earning its keep?"
if (!nrow(src_daily)) {
  cat("No source accounting yet — run rebuild_source_daily().")
} else {
  sc <- src_daily |> group_by(source_id) |>
    summarise(days = n_distinct(day), posts = sum(n_posts),
              coded = sum(n_coded), local = sum(n_local),
              pct_local = ifelse(sum(n_coded) > 0, sum(n_local)/sum(n_coded), NA_real_),
              mean_hhi = mean(hhi, na.rm = TRUE),
              novel = sum(novel_issues), .groups = "drop") |>
    mutate(local_per_day = round(local / pmax(days, 1), 2),
           pct_local = percent(pct_local, accuracy = 1),
           mean_hhi = round(mean_hhi, 2)) |>
    left_join(select(sources_tbl, source_id, platform, access), by = "source_id") |>
    arrange(desc(local_per_day)) |>
    select(Source = source_id, Platform = platform, Access = access,
           Days = days, Posts = posts, Coded = coded,
           `In-scope/day` = local_per_day, `% in scope` = pct_local,
           Concentration = mean_hhi, `Novel issues` = novel)
  knitr::kable(sc)
}
```

## Row {height="50%"}

```{r}
#| title: "Source registry and access posture"
if (!nrow(sources_tbl)) {
  cat("No source registry.")
} else {
  knitr::kable(sources_tbl |>
    mutate(Reply = ifelse(has_reply_edges, "yes", "no"),
           Active = ifelse(active, "yes", "no")) |>
    select(Source = source_id, Name = name, Platform = platform,
           Access = access, Cost = cost_note, `Reply edges` = Reply, Active))
}
```

::: {.provenance}
Sources are not interchangeable. Only platforms with stable pseudonymous IDs
and resolvable reply structure (`Reply edges = yes`) can support network
analysis at all; the rest contribute volume, issue mix and sentiment only.
**`In-scope/day`** — locally actionable comments per active day — is the honest
headline, because it rewards a source for producing material a council can act
on rather than for sheer volume. Access postures were verified on the dates
recorded in the registry and should be re-checked before any new crawl.
:::

# Network

## Row {height="52%"}

```{r}
#| title: "Reply network — who answers whom"
#| fig-width: 7.5
#| fig-height: 5
# SNA plotting practice applied here:
#  * layout seeded, so the figure is reproducible across renders
#  * node SIZE encodes degree, node COLOUR encodes community — two variables,
#    two channels, stated in the legend rather than left to be guessed
#  * edges drawn under nodes at low alpha so structure reads through hubs
#  * NO node labels: they are hashed ids that identify nobody and would only
#    add clutter
#  * n, m and density are in the caption, because a network picture without
#    them is decorative rather than evidential
if (!nrow(edges)) {
  cat("No reply edges yet.")
} else {
  suppressMessages({library(igraph); library(ggraph); library(tidygraph)})
  el <- edges |> filter(!is.na(from_user_id), !is.na(to_user_id),
                        from_user_id != to_user_id) |>
    count(from_user_id, to_user_id, name = "weight") |>
    mutate(across(1:2, as.character))
  g <- graph_from_data_frame(el, directed = TRUE)
  comp <- components(as_undirected(g, mode = "collapse"))
  g <- induced_subgraph(g, which(comp$membership == which.max(comp$csize)))
  set.seed(42)
  # Louvain is defined for UNDIRECTED graphs only — group_louvain() on a
  # directed tbl_graph errors out. Detect communities on an undirected copy,
  # then attach the membership back to the directed graph, which is what the
  # layout and the arrowheads need.
  memb <- membership(cluster_louvain(as_undirected(g, mode = "collapse")))
  tg <- as_tbl_graph(g) |>
    mutate(deg = centrality_degree(mode = "all"),
           community = as.factor(unname(memb[name])))
  ggraph(tg, layout = "fr") +
    geom_edge_link(aes(width = weight), alpha = 0.18, colour = BRAND$ink,
                   show.legend = FALSE) +
    scale_edge_width(range = c(0.2, 1.1)) +
    geom_node_point(aes(size = deg, colour = community), alpha = 0.88) +
    scale_size_continuous(range = c(1.6, 8), name = "Replies (degree)") +
    scale_colour_brewer(palette = "Dark2", guide = "none") +
    labs(title = "Reply network, largest connected component",
         subtitle = "Colour = community (Louvain); size = number of replies sent or received",
         # Caption lines must be short enough not to clip at the panel edge —
         # the first version ran off the right and truncated mid-sentence.
         caption = cap(sprintf(
           "%d actors, %d directed ties, density %.4f.\nAn edge is a quote-reply: A quoted B's specific post.\nStaff accounts and self-quotes excluded; node ids are hashes, not shown.",
           gorder(tg), gsize(tg), edge_density(tg)))) +
    theme_graph(base_family = "") +
    theme(text = element_text(size = 13, colour = BRAND$ink),
          plot.title = element_text(size = 16, face = "bold"),
          plot.subtitle = element_text(size = 12, colour = BRAND$muted),
          plot.caption = element_text(size = 12, colour = BRAND$muted, hjust = 0),
          legend.text = element_text(size = 12), legend.title = element_text(size = 12),
          legend.position = "right")
}
```

```{r}
#| title: "Degree distribution"
#| fig-width: 5
#| fig-height: 5
if (!nrow(edges)) {
  cat("No reply edges yet.")
} else {
  suppressMessages(library(igraph))
  el <- edges |> filter(from_user_id != to_user_id) |>
    count(from_user_id, to_user_id, name = "weight") |> mutate(across(1:2, as.character))
  g <- graph_from_data_frame(el, directed = TRUE)
  d <- data.frame(deg = degree(g, mode = "all"))
  ggplot(d, aes(deg)) +
    geom_histogram(binwidth = 1, fill = BRAND$accent, colour = "white", linewidth = 0.2) +
    scale_x_continuous(breaks = scales::pretty_breaks()) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.08))) +
    labs(x = "Degree (replies sent + received)", y = "Number of actors",
         title = "Participation is heavily skewed",
         caption = cap(sprintf("Median %d, max %d. A long tail of one-reply actors with a small core of hubs\nis the expected shape; it is why volume is always reported next to concentration.",
                           median(d$deg), max(d$deg)))) +
    theme_okcp(13)
}
```

## Row {height="48%"}

```{r}
#| title: "Structure of both networks"
if (!nrow(edges)) {
  cat("No reply edges yet.")
} else {
  suppressMessages(library(igraph))
  # Reply network
  el <- edges |> filter(from_user_id != to_user_id) |>
    count(from_user_id, to_user_id, name = "weight") |> mutate(across(1:2, as.character))
  gr <- graph_from_data_frame(el, directed = TRUE)
  # Co-participation: actors who posted in the same thread
  cop <- posts_meta |> filter(!is.na(actor_key)) |> distinct(actor_key, thread_t)
  gc <- NULL
  if (nrow(cop) > 10) {
    bi <- graph_from_data_frame(mutate(cop, thread_t = paste0("t", thread_t)), directed = FALSE)
    V(bi)$type <- startsWith(V(bi)$name, "t")
    pr <- bipartite_projection(bi, which = "false")
    gc <- subgraph_from_edges(pr, which(E(pr)$weight >= 2), delete.vertices = TRUE)
  }
  desc <- function(g) {
    if (is.null(g) || gorder(g) == 0) return(rep(NA, 6))
    gg <- if (is_directed(g)) as_undirected(g, mode = "collapse") else g
    cm <- components(gg); gt <- induced_subgraph(gg, which(cm$membership == which.max(cm$csize)))
    c(gorder(g), gsize(g), round(edge_density(g), 4), cm$no,
      round(transitivity(gt, "global"), 3), round(mean_distance(gt), 2))
  }
  tab <- data.frame(
    Measure = c("Actors", "Ties", "Density", "Components", "Transitivity", "Mean distance"),
    `Reply (directed)` = desc(gr), `Co-participation (>=2 threads)` = desc(gc),
    check.names = FALSE)
  knitr::kable(tab)
}
```

```{r}
#| title: "Small-world test"
if (!nrow(edges)) {
  cat("No reply edges yet.")
} else {
  suppressMessages(library(igraph))
  el <- edges |> filter(from_user_id != to_user_id) |>
    count(from_user_id, to_user_id, name = "weight") |> mutate(across(1:2, as.character))
  g <- as_undirected(graph_from_data_frame(el, directed = TRUE), mode = "collapse")
  cm <- components(g); gt <- induced_subgraph(g, which(cm$membership == which.max(cm$csize)))
  set.seed(42)
  sims <- replicate(50, {
    r <- sample_gnm(gorder(gt), gsize(gt))
    c(C = transitivity(r, "global"), L = mean_distance(r, unconnected = TRUE))
  })
  Co <- transitivity(gt, "global"); Lo <- mean_distance(gt)
  Cr <- mean(sims["C", ], na.rm = TRUE); Lr <- mean(sims["L", ], na.rm = TRUE)
  knitr::kable(data.frame(
    Measure = c("Clustering C (observed)", "Clustering C (random)",
                "Path length L (observed)", "Path length L (random)",
                "Small-world sigma"),
    Value = c(round(Co, 3), round(Cr, 3), round(Lo, 2), round(Lr, 2),
              round((Co / Cr) / (Lo / Lr), 2))))
}
```

::: {.provenance}
**Sigma above 1 indicates small-world structure** — clustering far above a
random graph of the same size and density, with comparable path lengths.
Compared against 50 Erdős–Rényi graphs, averaged over reachable pairs because
random graphs at this density are frequently disconnected.

**Read both networks together.** Quote-replies are the only unambiguous reply
signal in a flat phpBB thread, so the reply graph is sparse by construction and
understates real interaction. Co-participation is denser but weaker evidence:
two people in the same thread may never have addressed each other. A finding
that holds in only one of the two should say so.

**A caution on communities.** Louvain optimises modularity and always returns a
partition — even for a random graph — and its resolution limit hides small
groups. On this corpus the split is substantially by *thread*, not by stable
faction. A statistically validated backbone of the co-participation graph
(disparity filter, α = 0.05) reduced 3,723 raw ties to **15**: nearly all
apparent co-participation is explained by how active each actor is, not by
genuine affinity. Treat community structure as descriptive, not as evidence of
organised blocs.

**ERGM is fitted weekly and is currently suppressed.** MCMC did not mix on this
graph, and `fit_ergm()` reports the failure rather than printing coefficients
from a degenerate model. At 231 ties that is the expected outcome, not a bug —
it needs a larger corpus before the model is identifiable.
:::

## Row {height="50%"}

```{r}
#| title: "Network over time — growth, cohesion and turnover"
#| fig-width: 7.5
#| fig-height: 4.2
# LONGITUDINAL view. A single snapshot cannot tell a stable community from one
# that reassembles from different people every month, which is exactly what
# the retention series below shows is happening here.
if (!nrow(net_tl)) {
  cat("No timeline yet — run inst/dev/03_weekly.R.")
} else {
  d <- net_tl |> filter(view == "per-window") |>
    mutate(window = as.Date(window)) |>
    select(window, Actors = actors, Ties = ties,
           `Density` = density, `Transitivity` = transitivity) |>
    tidyr::pivot_longer(-window, names_to = "measure", values_to = "value") |>
    mutate(measure = factor(measure, c("Actors","Ties","Density","Transitivity")))
  ggplot(d, aes(window, value)) +
    geom_line(colour = BRAND$accent, linewidth = 0.9) +
    geom_point(colour = BRAND$accent, size = 2.2) +
    facet_wrap(~measure, scales = "free_y", nrow = 2) +
    scale_x_date(date_labels = "%b %Y") +
    labs(x = NULL, y = NULL,
         title = "The network is growing and thinning at the same time",
         caption = cap("Per-window (not cumulative). Density falls as actors join — expected. Falling transitivity\nmeans new participants are not closing triangles: the network is widening, not consolidating.")) +
    theme_okcp(13) + theme(strip.text = element_text(size = 13, face = "bold"))
}
```

```{r}
#| title: "Actor retention month to month"
#| fig-width: 5
#| fig-height: 4.2
if (!nrow(turnover)) {
  cat("No turnover data yet.")
} else {
  d <- turnover |> mutate(window = as.Date(window))
  ggplot(d, aes(window, retention)) +
    geom_col(fill = BRAND$accent, width = 20) +
    geom_text(aes(label = scales::percent(retention, accuracy = 1)),
              vjust = -0.4, size = pt(12), colour = BRAND$ink) +
    scale_y_continuous(labels = scales::percent,
                       expand = expansion(mult = c(0, 0.22)), limits = c(0, 1)) +
    scale_x_date(date_labels = "%b %Y") +
    labs(x = NULL, y = "Share of last month's actors still posting",
         title = "This is a revolving cast, not a stable community",
         caption = cap("Most participants each month are new. Treat 'community structure'\nas thread-level co-presence, not organised factions.")) +
    theme_okcp(13)
}
```

## Row {height="50%"}

```{r}
#| title: "ERGM — what predicts a reply tie"
#| fig-width: 6
#| fig-height: 3.6
# Coefficient plot with 95% CIs. An ERGM table of numbers is unreadable; the
# plot makes the one substantive result immediately legible.
if (!nrow(ergm_fit) || all(is.na(ergm_fit$estimate))) {
  cat(if (nrow(ergm_fit)) paste("ERGM suppressed:", ergm_fit$reason[1]) else "No ERGM fit yet.")
} else {
  d <- ergm_fit |> filter(!is.na(estimate)) |>
    mutate(lo = estimate - 1.96 * std_error, hi = estimate + 1.96 * std_error,
           sig = p_value < 0.05)
  ggplot(d, aes(estimate, reorder(term, estimate))) +
    geom_vline(xintercept = 0, linetype = "22", colour = BRAND$muted) +
    geom_errorbarh(aes(xmin = lo, xmax = hi), height = 0.14, colour = BRAND$ink) +
    geom_point(aes(colour = sig), size = 3.4) +
    scale_colour_manual(values = c("TRUE" = BRAND$accent, "FALSE" = BRAND$muted),
                        guide = "none") +
    labs(x = "Log-odds estimate (95% CI)", y = NULL,
         title = sprintf("Reciprocity dominates (spec: %s)", d$spec[1]),
         caption = cap(sprintf("n = %d actors, %d ties. Negative `edges` = the network is sparse.\nA positive `mutual` means a tie is far more likely when the reverse tie exists.",
                           d$n[1], d$m[1]))) +
    theme_okcp(13)
}
```

```{r}
#| title: "Small-world: observed vs random"
#| fig-width: 6
#| fig-height: 3.6
if (!nrow(edges)) {
  cat("No reply edges yet.")
} else {
  suppressMessages(library(igraph))
  el <- edges |> filter(from_user_id != to_user_id) |>
    count(from_user_id, to_user_id, name = "w") |> mutate(across(1:2, as.character))
  g <- as_undirected(graph_from_data_frame(el, directed = TRUE), mode = "collapse")
  cm <- components(g); gt <- induced_subgraph(g, which(cm$membership == which.max(cm$csize)))
  set.seed(42)
  sims <- replicate(50, {
    r <- sample_gnm(gorder(gt), gsize(gt))
    c(C = transitivity(r, "global"), L = mean_distance(r, unconnected = TRUE)) })
  d <- data.frame(
    measure = rep(c("Clustering (C)", "Path length (L)"), each = 2),
    which   = rep(c("Observed", "Random"), 2),
    value   = c(transitivity(gt, "global"), mean(sims["C", ], na.rm = TRUE),
                mean_distance(gt),          mean(sims["L", ], na.rm = TRUE)))
  sigma <- (d$value[1] / d$value[2]) / (d$value[3] / d$value[4])
  ggplot(d, aes(which, value, fill = which)) +
    geom_col(width = 0.62) +
    geom_text(aes(label = round(value, 3)), vjust = -0.35, size = pt(12), colour = BRAND$ink) +
    facet_wrap(~measure, scales = "free_y") +
    scale_fill_manual(values = c(Observed = BRAND$accent, Random = BRAND$muted), guide = "none") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.22))) +
    labs(x = NULL, y = NULL,
         title = sprintf("Small-world sigma = %.2f", sigma),
         caption = cap("Clustering far above a random graph of matched size and density, with comparable\npath lengths. Sigma > 1 indicates small-world structure. 50 Erdos-Renyi baselines.")) +
    theme_okcp(13) + theme(strip.text = element_text(size = 13, face = "bold"))
}
```

# Action

## Row {height="70%"}

```{r}
#| title: "Key take-home points"
#| output: asis
# GENERATED from the data on every render — see R/findings.R. Hardcoded
# findings are true the day they are written and quietly false afterwards.
# Every point carries its evidence and a stated confidence, because a
# dashboard that never reports its own limits invites over-reading.
# The error message MUST be surfaced, not swallowed. This tryCatch previously
# collapsed any failure into "not enough coded data", and when generate_findings()
# gained a `forums` argument the dashboard had never loaded, the entire Action
# tab quietly degraded to that sentence on a corpus of 35,462 coded rows. A
# code error and an empty corpus are different problems and must not print the
# same thing.
ferr <- NULL
f <- tryCatch(generate_findings(post_issues, posts_meta, issue_daily, sentiment,
                                src_daily, edges, forums, max_n = 10L),
              error = function(e) { ferr <<- conditionMessage(e); NULL })
if (!is.null(ferr)) {
  cat("**Findings could not be generated — this is a fault, not an empty corpus.**\n\n",
      "`generate_findings()` failed with: `", ferr, "`\n", sep = "")
} else if (is.null(f) || !nrow(f)) {
  cat("_Not enough coded data to generate findings._\n")
} else {
  badge <- c(strong = "1F7A8C", provisional = "B08900", weak = "C05621")
  for (i in seq_len(nrow(f))) {
    col <- badge[[f$confidence[i]]]
    cat(sprintf(
"::: {style=\"border-left:4px solid #%s; padding:6px 0 6px 12px; margin-bottom:14px;\"}
**%d. %s**  <span style=\"background:#%s;color:white;padding:1px 7px;border-radius:9px;font-size:11px;text-transform:uppercase;\">%s</span>

%s

*So what:* %s

<span style=\"color:#6B7A82;font-size:12.5px;\">Confidence — %s</span>
:::\n\n",
      col, i, f$headline[i], col, f$confidence[i],
      f$evidence[i], f$action[i], f$why[i]))
  }
}
```

## Row {height="30%"}

::: {.card title="Methods"}

**Corpus.** Public comments from Castanet's phpBB forums (nine forums, Okanagan
and BC) and agenda items plus public-hearing testimony from City of Kelowna
council records via eScribe. Every row records its `source_id`, so any figure
can be decomposed by source.

**Coding.** Each comment is coded against `codebook/codebook.yml` — a
content-hashed, versioned codebook of BC local-government competencies drawn
from the Community Charter and Local Government Act. The central variable is
**scope**: which order of government owns the issue. `ballot` is separate from
`local` because school trustees and regional district electoral area directors
are not municipal but appear on the same 2026-10-17 ballot. Every coded row
stores the codebook version that produced it, so revising a definition never
silently rewrites already-coded history.

**Classifier.** `claude-opus-5`, with structured outputs and a versioned
prompt. A keyword sieve runs first but is a **recall filter, not a finding** —
every figure here excludes sieve rows. An earlier cheaper tier was rejected on
measurement: it scored kappa 0.39 against Opus 5 on the scope judgement and
systematically under-called "local". Where multiple coders exist for a post,
strict precedence applies (human > Opus 5 > Sonnet 5 > Haiku) so figures never
average classifiers of different quality.

**Sentiment.** `sentimentr` polarity plus NRC emotion, **normalised per 100
words**. Raw NRC counts scale with comment length and produced a reversed
result before normalisation. Lexicon polarity failed to separate model-coded
stance (+0.03 oppose vs +0.09 support) and is therefore not reported as a
finding.

**Networks.** Two, deliberately. The *reply* network is directed and built from
blockquote citations, which resolve both the quoted user and the exact quoted
post — unambiguous but sparse. The *co-participation* network is denser but
weaker evidence. A disparity-filter backbone reduced 3,723 co-participation
ties to 15, indicating that nearly all apparent co-participation reflects
activity level rather than affinity. ERGM is fitted weekly over a ladder of
specifications and suppressed if none converges.

:::

::: {.card title="Limitations"}

- **Pilot scale.** Hundreds of items over an uneven window weighted to summer,
  when council agendas are thin.
- **Classifier agreement against a human standard is not yet measured.** The
  audit harness exists and a stratified sample is drawn; the figure is pending.
- **Discourse is not opinion.** Forum commenters are not a representative
  sample of residents. Volume measures attention, not public support.
- **Post-moderation corpus.** Removed comments and closed comment sections are
  invisible.
- **Council coverage is partial.** Matters reaching council through staff
  reports rather than numbered agenda items are not captured, so a zero in the
  agenda column may mean "not in this view" rather than "ignored".
- **Networks understate interaction.** Quoting is the only reply signal
  available in a flat phpBB thread.

:::

# Library

```{r}
#| title: "Report archive"
# Reports are generated artifacts, not tracked in git, so this indexes whatever
# exists on disk rather than assuming a fixed set.
rd <- file.path("..", "output", "reports")
files <- if (dir.exists(rd)) list.files(rd, pattern = "\\.pdf$", full.names = TRUE) else character()
if (!length(files)) {
  cat("No reports generated yet.\n\nRun `Rscript inst/dev/06_daily_report.R` to produce one.")
} else {
  info <- file.info(files)
  d <- data.frame(
    file = basename(files),
    date = as.Date(sub(".*?(\\d{4}-\\d{2}-\\d{2})\\.pdf$", "\\1", basename(files))),
    size = paste0(round(info$size / 1024), " KB"),
    stringsAsFactors = FALSE)
  d <- d[order(d$date, decreasing = TRUE), ]
  d$Year <- format(d$date, "%Y"); d$Month <- format(d$date, "%B")
  d$Report <- sprintf("[%s](../output/reports/%s)", d$file, d$file)
  knitr::kable(d[, c("Year", "Month", "date", "Report", "size")],
               col.names = c("Year", "Month", "Date", "Report", "Size"),
               row.names = FALSE)
}
```

::: {.provenance}
Briefs are **generated artifacts** and are not committed to the repository —
they regenerate from `report/daily.qmd` against the current corpus. The daily
driver refuses to render if the corpus is more than three days stale, so a
brief dated today is never built from week-old data.

Every brief is de-texted: aggregate counts, issue codes and network structure
only. No comment text, no handles, no speaker names.
:::
